---
title: "MIMIC-III ADMISSIONS TABLE"
output: "html_notebook"
editor_options: 
  chunk_output_type: inline
---

# LOS ICU Range definition
```{r}
X<-1.8
Y<-10
```


## Libraries
```{r library, message=FALSE}
library(tidyverse)
library(ggfittext)
library(skimr)
library(funModeling)
library(magrittr)
```

---

## Loading the Datasets
```{r loading data, message=T, warning=T}
#setwd("~/Documents/_DATA/original/")
setwd("~/Documents/Documents/THESIS/THESIS-II/thesis/NOTEBOOKS/_DATA/original")
df_adm  <- read_csv("ADMISSIONS.csv")
```

___
---

## Exploratory Data Analysis

```{r, eval = FALSE}
skim(df_adm)
```


```{r}
df_status(df_adm)
```

QUALITATIVE VARIABLES

ADMISSION_TYPE
```{r bar plot, eval=FALSE}
ggplot(data = df_adm, aes(x = ADMISSION_TYPE)) +
  geom_bar(color="white", fill="steelblue") +
  geom_text(stat='count', aes(label=..count..), vjust=-0.4)+
  ylim(-1000, 45000)
```

ADMISSION_LOCATION
```{r bar plot, eval=FALSE}
Location<-fct_infreq(df_adm$ADMISSION_LOCATION)
ggplot(data = df_adm, aes(x = Location)) +
  geom_bar(color="white", fill="steelblue") +
  geom_text(stat='count', aes(label=..count..), hjust=-0.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0, 25000) +
    coord_flip()
```

DISCHARGE_LOCATION
```{r bar plot discharge loc, eval=FALSE}
Discharge_location <- fct_infreq(df_adm$DISCHARGE_LOCATION)
df_adm %>% 
ggplot(aes(x = Discharge_location)) +
  geom_bar(color="white", fill="steelblue") +
  geom_text(stat='count', aes(label=..count..),hjust = -0.25 , vjust = 0.5, size = 3.5) +
  theme(axis.text.x = element_text(angle = 0,  hjust = 0.5   , vjust = 0  , size = 10), 
        axis.text.y = element_text(angle = 0,  hjust = 0     , vjust = 0  , size = 10)) +
  ylim(0, 22000)+
  coord_flip()
```

```{r, eval = FALSE}
df_adm %>% 
  select(DISCHARGE_LOCATION) %>% 
  unique() #>% 
  #arrange()
```


INSURANCE
```{r Insurance, eval=FALSE}
Insurance <- fct_infreq(df_adm$INSURANCE)
ggplot(data = df_adm, aes(x = Insurance)) +
  geom_bar(color="white", fill="steelblue") +
  geom_text(stat='count', aes(label=..count..), hjust=-0.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 12)) +
  ylim(0, 32000) +
  coord_flip()
```

MARITAL_STATUS
```{r Marital Status, eval=FALSE}
marital_status<- fct_infreq(df_adm$MARITAL_STATUS)
ggplot(data = df_adm, aes(x = marital_status)) +
  geom_bar(color="white", fill="steelblue") +
  geom_text( stat='count', aes(label=..count..), hjust=-0.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0, 30000)+
  coord_flip()
```

RELIGION
```{r Religion, eval=FALSE}
religion<- fct_infreq(df_adm$RELIGION)
ggplot(data = df_adm, aes(x = religion)) +
  geom_bar(color="white", fill="steelblue") +
  geom_text( stat='count', aes(label=..count..), hjust=-0.3) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  ylim(0, 30000)+
  coord_flip()
```

LANGUAGE

```{r Language}
df_adm$LANGUAGE %>% unique()

```

ETHNICITY
```{r Ethnicity, fig.height = 20, fig.width = 20, eval=FALSE}
df_adm %>% 
  select(ETHNICITY) %>% 
ggplot(aes(x = fct_infreq(ETHNICITY))) +
  geom_bar(width = 0.95, color="white", fill="steelblue") +
  geom_text(size = 6, stat='count', aes(label=..count..), hjust=-0.3) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.4, size =16), axis.text.y = element_text(size =16)) +
  ylim(0, 44000) +
  coord_flip()
```


#### Cheking unique patients and admissions.

Number of Unique patients
```{r}
pat_n <- df_adm %>% 
  distinct(SUBJECT_ID) %>% 
  count()
print(paste("Number of unique Patients: ", pat_n))
```

Number of Unique admissions
```{r}
ad_n <- df_adm %>% 
  distinct(HADM_ID) %>% 
  count()
print(paste("Number of unique Admissions: ", ad_n))
```

Readmission rate
```{r}
readim <- round((ad_n/pat_n - 1), 4)*100
print(paste("Rate of readmissions:", readim, "%"))
```


--- A
#### Removing the obvious Newborn patients which are not the demographic of interest (Age >50)

Checking number of Admission Type
```{r}
df_adm$ADMISSION_TYPE %>% unique()
```

Counting the number of Newborn Admissions
```{r}
NB_adm_type <- df_adm %>% 
  filter(ADMISSION_TYPE == "NEWBORN") %>% count()
print(paste("Newborn Admissions: ", NB_adm_type))
```

```{r removing newborn}
df_adm_A <- df_adm %>% 
  filter(ADMISSION_TYPE != "NEWBORN")
```

--- B

#### Removing Admissions with result in DEATH.
If the admissions resulting in DEATH should not be included because the LOS is consider that
the patient left alive from the hospital.
I can use the `HOSPITAL_EXPIRATION_FLAG` to filter such admissions.

```{r}
df_adm_A %>% 
  filter(HOSPITAL_EXPIRE_FLAG == 1)
```

```{r}
df_adm_A %>% 
  filter(HOSPITAL_EXPIRE_FLAG == 1) %>% 
  count() 
```

```{r filtering expired patient in ICU}
df_adm_B <- df_adm_A %>% 
  filter(!HOSPITAL_EXPIRE_FLAG == 1) 
```

--- C

#### Calculating the LOS and Removing the Admissions with negative LOS.
Negative LOS indicates that the patient died before the admission.`
```{r}
df_adm_B$LOS_HOSP <- round(as.numeric(difftime(df_adm_B$DISCHTIME, df_adm_B$ADMITTIME, units='days')),2)
```

Checking the statistical description of the new feature.
```{r, eval = FALSE}
describe(df_adm_B$LOS_HOSP)
```

Checking the negative and positive LOS_HOSP
```{r}
neg_los <- df_adm_B %>% 
  filter(LOS_HOSP < 0) %>% 
  nrow()
print(paste("Number of admissions with negative LOS: ", neg_los))
```

```{r}
pos_los <- df_adm_B %>% 
  filter(LOS_HOSP > 0) %>% 
  nrow()
print(paste("Number of admissions with positive LOS: ", pos_los))
```

Dropping negative LOS_HOSP
```{r dropping negative LOS}
df_adm_C <- df_adm_B %>% 
  filter(LOS_HOSP > 0)
```

```{r, eval = FALSE}
df_adm_B$LOS_HOSP %>%
  describe()
```

Checking the result searching for unique admissions
```{r, eval = FALSE}
df_adm_C %>% 
  n_distinct(vars(HADM_ID)) 
```
Descriptive analysis after the drop negative values.
```{r, eval = FALSE}
describe(df_adm_C$LOS_HOSP)
```


```{r, warning= FALSE, message=FALSE, eval = FALSE}
df_adm_A %>% describe()
df_adm_B %>% describe()
df_adm_C %>% describe()
```

#### Calculating the `AGE`

Importing the Patient table (already cleaned)
```{r Importing df_pat_x}
#setwd("~/Documents/_DATA/subset/")
setwd("~/Documents/Documents/THESIS/THESIS-II/thesis/NOTEBOOKS/_DATA/subset")
df_pat_A <- read_csv("df_pat_5.csv")
```

--- D

Inner join `Patient` (A) with `Admission` (C)
It is necessary to add DOB to the admission to calculate the `AGE`.

```{r Inner join Patient and Admission}
df_adm_D <- df_adm_C %>% 
  inner_join ( df_pat_A , by = "SUBJECT_ID" ) 
```

```{r, warning=FALSE, message = FALSE, eval=FALSE}
df_adm_D %>%  describe() 
```

```{r Creating the variable AGE}
df_adm_D$ADMITTIME <- as.Date(df_adm_D$ADMITTIME)
df_adm_D$DOB <- as.Date(df_adm_D$DOB)

df_adm_D$AGE <- round(
  as.numeric(difftime(df_adm_D$ADMITTIME, df_adm_D$DOB, 
                      units = "days")/365.242), 2)
```

Filtering the Admissions entries for patients with `AGE` < 50, which are part of demograpgic of interest.
It was filtered all the patients with age greater than 89 years old because such patients had their
DOB anonimized preventing of calculate the correct age.
(https://mimic.physionet.org/mimicdata/time/)


```{r, eval = FALSE}
colnames(df_adm_D)
```


```{r histogram plot, eval=FALSE}
ggplot(df_adm_D, aes(x=AGE, fill=GENDER, color=GENDER)) +
  geom_histogram(binwidth = 1, position="identity", alpha=0.3)

```



```{r}
df_adm_D %>% 
  filter(AGE > 90) %>% 
  count()
```


```{r Filtering AGE 50~90}
df_adm_D %<>% 
  filter(AGE >50 & AGE <90)
```

```{r, eval=FALSE}
df_adm_D %>% 
  ggplot(aes(x = AGE))+
  geom_histogram(binwidth = 1,  color="white", fill="steelblue")

```

Once I have the dataset filtered with the AGE of interest some Feature Transformation are necessary.

```{r, eval=FALSE}
df_adm_D %>% 
 # filter(LOS_HOSP < 12) %>% 
ggplot(aes(x = LOS_HOSP))+
geom_histogram(binwidth = 2,  color="white", fill="steelblue")+
  xlim(0, 60)
```

ETHNICITY
```{r fig.height = 20, fig.width = 20, eval=FALSE}
df_adm %>% 
  select(ETHNICITY) %>% 
ggplot(aes(x = fct_infreq(ETHNICITY))) +
  geom_bar(width = 0.95, color="white", fill="steelblue") +
  geom_text(size = 6, stat='count', aes(label=..count..), hjust=-0.3) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.4, size =16), axis.text.y = element_text(size =16)) +
  ylim(0, 44000) +
  coord_flip()
```

```{r, eval = FALSE}
df_adm_D$LOS_HOSP %>%
  describe()
```

#### Ethnicity

```{r fig.height = 20, fig.width = 20, eval=FALSE}

df_adm_D %>% 
  ggplot(aes(x = fct_infreq(ETHNICITY), y = LOS_HOSP))+
  geom_boxplot(color="black", fill="steelblue") +
  theme(
    axis.title.x = element_text(size = 24),
    axis.text.x = element_text(angle = 0, hjust = 0.4, size =16), 
    axis.text.y = element_text(size =16),
    axis.title.y = element_text(size = 24)
    ) +
  labs(x="Ethnicity", y = "Count") +
  coord_flip()

```

It is necessary to group the several ethnicities variations. 
Ethnicities:

1. White
2. Latino
3. Black
4. Asian
5. Other

```{r}
eth_list <- df_adm_D %>%
   count(ETHNICITY) %>% 
  arrange(desc(n)) %>% 
  as_tibble() 

eth_list
```

```{r}
a <- round((eth_list$n/sum(eth_list$n)*100),2)
eth_list$n_perc <- a
eth_list
```


```{r}
eth_list$n_perc <- eth_list[,3]
eth_list$n_perc <- eth_list[,-3]
eth_list
# write_csv(eth_list, "data/ethnical_groups.csv")
```


1. White: WHITE, PORTUGUESE, "WHITE - OTHER EUROPEAN", "MIDDLE EASTERN", "WHITE - EASTERN EUROPEAN"

2. Latino: "HISPANIC OR LATINO", "SOUTH AMERICAN", "HISPANIC/LATINO - PUERTO RICAN", "WHITE - BRAZILIAN", "HISPANIC/LATINO - GUATEMALAN", "HISPANIC/LATINO - DOMINICAN", "HISPANIC/LATINO - CENTRAL AMERICAN (OTHER)", "HISPANIC/LATINO - MEXICAN", "HISPANIC/LATINO - SALVADORAN"

3. Black: "BLACK/AFRICAN AMERICAN", "BLACK/CAPE VERDEAN", "BLACK/HAITIAN", "BLACK/AFRICAN", 

4. Asian: ASIAN, "ASIAN - CHINESE", "ASIAN - JAPANESE", "ASIAN - VIETNAMESE", "ASIAN - OTHER", "ASIAN - ASIAN INDIAN", "ASIAN - KOREAN", "ASIAN - FILIPINO", 
5. Other: OTHER, "MULTI RACE ETHNICITY", "AMERICAN INDIAN/ALASKA NATIVE", "NATIVE HAWAIIAN OR OTHER PACIFIC ISLANDER", "AMERICAN INDIAN/ALASKA NATIVE FEDERALLY RECOGNIZED TRIBE"

6. Unkown: "UNKNOWN/NOT SPECIFIED", "PATIENT DECLINED TO ANSWER",  "UNABLE TO OBTAIN", 


```{r Mapping ethinicity}
df_adm_D$ETHNICITY <- as.character(df_adm_D$ETHNICITY)
df_adm_D <- df_adm_D %>%
  mutate(ETHNICITY_MAPPED = case_when(
    
    # Other
    ETHNICITY == "OTHER" ~ "Other",
    ETHNICITY == "MULTI RACE ETHNICITY" ~ "Other",
    ETHNICITY == "AMERICAN INDIAN/ALASKA NATIVE" ~ "Other",
    ETHNICITY == "NATIVE HAWAIIAN OR OTHER PACIFIC ISLANDER" ~ "Other",
    ETHNICITY == "MIDDLE EASTERN" ~ "Other",
    ETHNICITY == "AMERICAN INDIAN/ALASKA NATIVE FEDERALLY RECOGNIZED TRIBE" ~ "Other",
    
    # Unkown
    ETHNICITY == "UNKNOWN/NOT SPECIFIED" ~ "Unkown",
    ETHNICITY == "PATIENT DECLINED TO ANSWER" ~ "Unkown",
    ETHNICITY == "UNABLE TO OBTAIN" ~ "Unkown",
    
    #Asian
    ETHNICITY == "ASIAN" ~ "Asian",
    ETHNICITY == "ASIAN - CHINESE" ~ "Asian",
    ETHNICITY == "ASIAN - JAPANESE" ~ "Asian",
    ETHNICITY == "ASIAN - VIETNAMESE" ~ "Asian",
    ETHNICITY == "ASIAN - OTHER" ~ "Asian",
    ETHNICITY == "ASIAN - ASIAN INDIAN" ~ "Asian",
    ETHNICITY == "ASIAN - KOREAN" ~ "Asian",
    ETHNICITY == "ASIAN - FILIPINO" ~ "Asian",
    ETHNICITY == "ASIAN - THAI" ~ "Asian",
    ETHNICITY == "ASIAN - CAMBODIAN" ~ "Asian",
    
    #Black
    ETHNICITY == "BLACK/AFRICAN AMERICAN" ~ "Black",
    ETHNICITY == "BLACK/CAPE VERDEAN" ~ "Black",
    ETHNICITY == "BLACK/HAITIAN" ~ "Black",
    ETHNICITY == "BLACK/AFRICAN" ~ "Black",
    
    #Latino
    ETHNICITY == "HISPANIC OR LATINO" ~ "Latino",
    ETHNICITY == "SOUTH AMERICAN" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - PUERTO RICAN" ~ "Latino",
    ETHNICITY == "WHITE - BRAZILIAN" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - GUATEMALAN" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - DOMINICAN" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - CENTRAL AMERICAN (OTHER)" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - MEXICAN" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - SALVADORAN" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - CUBAN" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - HONDURAN" ~ "Latino",
    ETHNICITY == "HISPANIC/LATINO - COLOMBIAN" ~ "Latino",
    ETHNICITY == "CARIBBEAN ISLAND" ~ "Latino",
    
    #White
    ETHNICITY == "WHITE" ~ "White",
    ETHNICITY == "PORTUGUESE" ~ "White",
    ETHNICITY == "WHITE - OTHER EUROPEAN" ~ "White",
    ETHNICITY == "WHITE - EASTERN EUROPEAN" ~ "White",
    ETHNICITY == "WHITE - RUSSIAN" ~ "White",
    TRUE ~ ETHNICITY
  ))
df_adm_D$ETHNICITY_MAPPED <- as.factor(df_adm_D$ETHNICITY_MAPPED)
```

```{r}
df_adm_D$ETHNICITY_MAPPED %>% unique()
```

```{r}
df_adm_D$ETHNICITY_MAPPED %>%
  skim()
```


ETHNICITY
```{r, eval=FALSE}
df_adm_D %>% 
ggplot(aes(x = (fct_infreq(ETHNICITY_MAPPED)))) +
  geom_bar(width = 0.95, color="white", fill="steelblue") +
  geom_text(size = 5, stat='count', aes(label=..count..), hjust=-0.3) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.4, size =14), axis.text.y = element_text(size =14)) +
  ylim(0, 28000) +
  xlab("Ethnicity") +
  coord_flip()
```

Boxplots to identify outliers
    
```{r, eval=FALSE}
df_adm_D %>% 
 # filter(LOS_HOSP < 12) %>% 
ggplot(aes(x = ETHNICITY_MAPPED, y = LOS_HOSP, fill=ETHNICITY_MAPPED))+
geom_boxplot(alpha = 0.7) +
    theme(legend.position="center", axis.text.x = element_text(angle = 0, hjust = 0.4, size =14), axis.text.y = element_text(size =14))+ 
  xlab("Ethnicity") +
  ylab("LOS - Hospital") +
    coord_flip()
```


Filtering the outliers
```{r, eval=FALSE}
df_adm_D %>% 
  filter(LOS_HOSP < 50) %>% 
ggplot(aes(x = ETHNICITY_MAPPED, y = LOS_HOSP, fill=ETHNICITY_MAPPED))+
geom_boxplot(alpha = 0.7) +
    theme(legend.position="center", axis.text.x = element_text(angle = 0, hjust = 0.4, size =14), axis.text.y = element_text(size =14))+ 
  xlab("Ethnicity") +
  ylab("LOS - Hospital") +
    coord_flip()
```

Counting the outliers
```{r}
n_out<- df_adm_D %>%
  filter(LOS_HOSP >50) %>% 
  count()
n_nor<- df_adm_D %>%
  filter(LOS_HOSP <=50) %>% 
  count()
a <- c("outliers", "norm")
b <- rbind(n_out, n_nor)  

cbind(a,b) %>% as_tibble()
```


Dropping variables.

```{r, eval = FALSE}
df_adm_D %>% 
  colnames()
```

```{r dropping variables}
df_adm_E <- df_adm_D %>% 
  select(HADM_ID, SUBJECT_ID, GENDER, AGE, ETHNICITY_MAPPED, ADMISSION_TYPE, LOS_HOSP)
```


```{r, eval = FALSE}
df_adm_E %>%  head()
```

Joinning with `df_stay_A`

Getting the default ICU_LOS

Loading df_stay_A.csv
```{r loading data stay}
#setwd("~/Documents/_DATA/subset/")
setwd("~/Documents/Documents/THESIS/THESIS-II/thesis/NOTEBOOKS/_DATA/subset")
df_stay_A <- read_csv("df_stay_5.csv")
```

```{r inner join stay}
df_adm_F <- df_adm_E %>% 
  inner_join ( df_stay_A , by = "HADM_ID" ) 
```

```{r}
df_adm_F %<>% select(-LOS_HOSP, LOS_HOSP)
df_adm_F %<>% rename(LOS_ICU = LOS)
```


Checking the LOS_ICU
```{r boxplot eth vs LOS-ICU, eval=FALSE}
df_adm_F %>% 
#  filter(LOS_ICU < 6) %>% 
ggplot(aes(x = ETHNICITY_MAPPED, y = LOS_ICU, fill=ETHNICITY_MAPPED))+
geom_boxplot(alpha = 0.7) +
    theme(legend.position="center", axis.text.x = element_text(angle = 0, hjust = 0.4, size =14), axis.text.y = element_text(size =14))+ 
  xlab("Ethnicity") +
  ylab("LOS-ICU") +
    coord_flip()
```

```{r}
df_adm_F %>% dim()
```

Filtering the outliers
```{r Filtering Outliers}
df_adm_F %<>% 
  filter(LOS_ICU < 100)
```


```{r LOS Diff calc}
df_adm_F %<>% 
  mutate(LOS_diff = LOS_HOSP -LOS_ICU) 
```


```{r, eval = FALSE}
df_adm_F %>% 
  filter(LOS_diff <=0) %>% 
  select(LOS_diff) %>% 
  summary()
```

Drop the admission which the LOS_ICU is greater than LOS_HOSP. Something is wrong with such data, how can the total stay in the hospital be smaller than ICU stay?
```{r}
df_adm_F %>% nrow()
df_adm_F %<>% 
  filter(LOS_diff>0)
df_adm_F %>% nrow()
```

```{r, eval = FALSE}
df_adm_F %>%  colnames()
```

Removing variables which it will not be used
```{r dropping variables}
df_adm_F %<>% 
  select(-SUBJECT_ID.y, -ICUSTAY_ID, 
         -LAST_CAREUNIT, 
         -INTIME, 
         -OUTTIME, 
         -LOS_diff)
```


```{r, eval = FALSE}
df_adm_F %>%  colnames()
```

Creating classification groups for LOS_ICU

```{r LOS ranges, eval = TRUE}
#X<-3
#Y<-10

df_adm_F %<>% 
  mutate(LOS_GROUP_ICU = case_when(
    LOS_ICU <= X ~ "A",
    LOS_ICU >X & LOS_ICU <Y ~ "B",
    LOS_ICU >= Y ~ "C",
    TRUE ~ "O"
  ))

df_adm_F %>% 
  select(LOS_GROUP_ICU) %>% 
  table() 
```

```{r}
df_adm_F$LOS_GROUP_ICU <- as.factor(df_adm_F$LOS_GROUP_ICU)
```

```{r}
df_adm_F %>% 
  select(LOS_GROUP_ICU) %>% 
  table() %>% 
  prop.table() %>% 
  round(3)
```

```{r, eval = FALSE}
df_adm_F %>% colnames()
```

```{r dropping variables}
df_adm_G <- df_adm_F %>%
  select(-LOS_ICU,   LOS_ICU, 
         -LOS_HOSP , 
         -LOS_GROUP_ICU, LOS_GROUP_ICU
         )
```

```{r}
df_adm_G %>% colnames()
```

```{r}
df_adm_G %<>% rename(SUBJECT_ID = SUBJECT_ID.x)
```

```{r, eval = FALSE}
df_adm_G %>% 
  glimpse()
```

```{r}
df_adm_G %<>% mutate_if(is_character, as_factor)
```

```{r exporting the final df}
setwd("~/Documents/_DATA/subset/")
write_csv(df_adm_G, "df_adm_1.8-10.csv")
```


