---
title: "MIMIC-III DIAGNOSES TABLE"
output: "html_notebook"
---
Test with ADM_H
has cut for outliers >15

## Libraries
```{r, message=FALSE}
library(tidyverse)
library(stringi)
library(stringr)
library(splitstackshape)
library(qdapTools)
library(psych)
```
____
## Loading the Datasets
```{r}
setwd("~/Documents/_DATA/original/")
df_diag <- read_csv("DIAGNOSES_ICD.csv")
```
___
## Exploratory Data Analysis

```{r}
df_diag %>% glimpse()
```


```{r, warning=FALSE, message=FALSE}
df_diag %>% describe()
```

Unique ICD9 Codes:

```{r, warning=FALSE, message=FALSE}
df_diag %>% 
  select(ICD9_CODE) %>%
  unique() %>% 
  count()
```
Filter the admissions resulted in the ADMISSIONS (E) notebook.

```{r}
setwd("~/Documents/_DATA/subset/")
df_adm_G <- read_csv("df_adm_5.csv")
```

Using the vector of HADM_ID from the ADMISSIONS df_adm_F.csv whic contains pattients older than 50 yo and younger than 89 yo.
```{r}
df_admin_G_col <- df_adm_G %>% 
  select(HADM_ID)

df_diag_A <- left_join(df_admin_G_col, df_diag)
df_diag_A
```

Unique ICD9 Codes after used for the admissions filtered in the ADMISSIONS notebook:

```{r, warning=FALSE, message=FALSE}
df_diag_A %>% 
  select(ICD9_CODE) %>% 
  unique() %>% 
  count()
```

```{r}
icd9_table <- df_diag_A %>% count(ICD9_CODE) %>% 
  arrange(desc(n) ) %>% 
  mutate(perc = round((n/sum(n))*100,3)) %>% 
  as_tibble() %>% 
    head(30)
icd9_table
```

```{r}
icd9_table %>% select(perc) %>% 
  sum()

#Top 30 ICD Code represents (%)
```


###Creating a new df to receive the new ICD9 mapping (df_diag_B).

```{r}
df_diag_B <- df_diag_A
```

Recode the ICD9 code to represent the code family generalizing the variable

ICD9 Codes Source: http://www.icd9data.com/2015/Volume1/default.htm

#### Mapping the IDC9 codes

Initiating new collumns
```{r}
df_diag_B$ICD9_FAMILY  <- ""
df_diag_B$ICD9_V       <- 0
df_diag_B$ICD9_E       <- 0
df_diag_B$ICD9_NUMBERS <- 0
```

#### Encoding ICD9 V codes

 V15.82: Personal history of tobacco use.
 V45.81: Aortocoronary bypass status.
 V45.82: Percutaneous transluminal coronary angioplasty status.
 V58.61: Long-term (current) use of anticoagulants.
 __Remaining codes:__
 V01~V91: Factors Influencing Health Status
 
```{r}
df_diag_B$ICD9_V <- ifelse((str_detect(df_diag_B$ICD9_CODE, regex("^V"))), 
                           sub(".", "", df_diag_B$ICD9_CODE), # Remove the first character of the string
                           0)
stri_sub(df_diag_B$ICD9_V, 3,2) <- "."
df_diag_B$ICD9_V <- df_diag_B$ICD9_V %>% as.numeric()
```

Dropping NA's in ICD9_V
```{r}
df_diag_B <- df_diag_B %>% 
  drop_na(ICD9_V)
```

Mapping `V` codes.

```{r}
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_V >=     1] <- "Factors_influencing"
#Now overwritting the previous string in specific cases.
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_V == 15.82] <- "Tobacco_use"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_V == 45.81] <- "Aortocoronary_bypass"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_V == 45.82] <- "Coronary_angioplasty"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_V == 58.61] <- "Anticoagulants"
```

Verifying the mapping process
```{r}
df_diag_B$ICD9_FAMILY %>%  
  unique()
```
The "" displayed in the chunk above is related to ICD codes diferent than "V".

#### Encoding ICD9 E codes

E000~E999:
E000.X~E030 "External"
E800~E807   "External"
E810~E848   "Accident"
E849        "Place_of_occurrence"
E850~E858   "Poisoning"
E860~E869   "Poisoning"
E870~E876   "Complication"
E878~E8879  "Complication"
E880~E888   "Accident"
E890~E899   "Accident"
E900~E909   "Accident"

`E` ICD9 codes of interest:
The top 30 codes does not have any code `E`. Based on this I will use the
families' names.

```{r}
df_diag_B$ICD9_E <- ifelse({str_detect(df_diag_B$ICD9_CODE, regex("^E", ignore_case = TRUE))}, sub(".", "", df_diag_B$ICD9_CODE), 0)
```

```{r}
df_diag_B$ICD9_E <- substr(df_diag_B$ICD9_E, start = 1, stop = 3)
```

```{r}
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >  0   & df_diag_B$ICD9_E <= 30]  <- "External"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 800 & df_diag_B$ICD9_E <= 848] <- "Accident"
# df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 840 & df_diag_B$ICD9_E <= 848] <- "Accident"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E == 849] <- "Place_of_occurrence"

df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 850 & df_diag_B$ICD9_E <= 869] <- "Poisoning"
# df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 860 & df_diag_B$ICD9_E <= 869] <- "Poisoning"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 870 & df_diag_B$ICD9_E <= 879] <- "Misadventure"
# df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 878 & df_diag_B$ICD9_E <= 879] <- "Misadventure"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 880 & df_diag_B$ICD9_E <= 929] <- "Accident"
# df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 890 & df_diag_B$ICD9_E <= 899] <- "Accident"
# df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 900 & df_diag_B$ICD9_E <= 929] <- "Accident"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 930 & df_diag_B$ICD9_E <= 949] <- "Adverse_Effect"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 950 & df_diag_B$ICD9_E <= 969] <- "Nonnatural"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 970 & df_diag_B$ICD9_E <= 979] <- "Legal"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 980 & df_diag_B$ICD9_E <= 989] <- "Undetermined"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_E >= 990 & df_diag_B$ICD9_E <= 999] <- "War"

```

Verifying the mapping process
```{r}
df_diag_B$ICD9_FAMILY %>% unique()
```

#### Encoding ICD9 Numerical codes

000~999:
  001-139  "Infectious" - Infectious And Parasitic Diseases
  140-239  "Neoplasms" - Neoplasms
  240-279  "Metabolic" - Endocrine, Nutritional And Metabolic Diseases, And Immunity Disorders
  280-289  "Blood" - Diseases Of The Blood And Blood-Forming Organs
  290-319  "Mental" - Mental Disorders
  320-389  "Nervous" - Diseases Of The Nervous System And Sense Organs
  390-459  "Circulatory" - Diseases Of The Circulatory System
  460-519  "Respiratory" - Diseases Of The Respiratory System
  520-579  "Digestive" - Diseases Of The Digestive System
  580-629  "Genitourinary" - Diseases Of The Genitourinary System
  630-679  "Pregnancy" - Complications Of Pregnancy, Childbirth, And The Puerperium
  680-709  "Skin" - Diseases Of The Skin And Subcutaneous Tissue
  710-739  "Musculoskeletal" - Diseases Of The Musculoskeletal System And Connective Tissue
  740-759  "Congenital" - Congenital Anomalies
  760-779  "Perinatal" - Certain Conditions Originating In The Perinatal Period
  780-799  "Ill-Defined" - Symptoms, Signs, And Ill-Defined Conditions
  800-999  "Injury_Poisoning" - Injury And Poisoning

```{r}
df_diag_B$ICD9_NUMBERS <- ifelse(
                     (str_detect(df_diag_B$ICD9_CODE, regex("^[^a-zA-Z]", ignore_case = TRUE)))
                     ,substr(df_diag_B$ICD9_CODE, start = 1, stop = 3), 0)
```

Mapping the ICD9 code to the correspondent family
```{r}
df_diag_B$ICD9_NUMBERS <- as.numeric(df_diag_B$ICD9_NUMBERS)
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >=   1 & df_diag_B$ICD9_NUMBERS < 140] <- "Infectious"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 140 & df_diag_B$ICD9_NUMBERS < 240] <- "Neoplasms"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 240 & df_diag_B$ICD9_NUMBERS < 280] <- "Metabolic"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 280 & df_diag_B$ICD9_NUMBERS < 290] <- "Blood"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 290 & df_diag_B$ICD9_NUMBERS < 320] <- "Mental"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 320 & df_diag_B$ICD9_NUMBERS < 390] <- "Nervous"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 390 & df_diag_B$ICD9_NUMBERS < 460] <- "Circulatory"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 460 & df_diag_B$ICD9_NUMBERS < 520] <- "Respiratory"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 520 & df_diag_B$ICD9_NUMBERS < 580] <- "Digestive"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 580 & df_diag_B$ICD9_NUMBERS < 630] <- "Genitourinary"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 630 & df_diag_B$ICD9_NUMBERS < 680] <- "Pregnancy"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 680 & df_diag_B$ICD9_NUMBERS < 710] <- "Skin"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 710 & df_diag_B$ICD9_NUMBERS < 740] <- "Musculoskeletal"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 740 & df_diag_B$ICD9_NUMBERS < 760] <- "Congenital"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 760 & df_diag_B$ICD9_NUMBERS < 780] <- "Perinatal"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 780 & df_diag_B$ICD9_NUMBERS < 800] <- "Ill-Defined"
df_diag_B$ICD9_FAMILY[df_diag_B$ICD9_NUMBERS >= 800 & df_diag_B$ICD9_NUMBERS <= 999]<- "Injury_Poisoning"

```

Verifying the mapping process
```{r}
df_diag_B$ICD9_FAMILY %>% unique()
```

```{r}
df_diag_B %>% count(ICD9_FAMILY) %>% 
  arrange(desc(n) ) %>% 
  mutate(perc = round((n/sum(n))*100,3)) %>% 
  as_tibble()

```

Let's drop auxiliar columns.
```{r}
df_diag_B %>% 
  colnames()
```
```{r}
df_diag_B %<>% 
  select(
    -ICD9_V, 
    -ICD9_E, 
    -ICD9_NUMBERS) 
```


#### Groupding the IDC9 Codes per `HADM_ID`
     
```{r}
df_diag_C <- df_diag_B %>% 
     group_by(HADM_ID) %>% 
     mutate(ICD9_CODES_LIST = paste0(ICD9_FAMILY, collapse = ",")) %>% 
     filter(SEQ_NUM == 1) %>%
     select(HADM_ID, SUBJECT_ID, ICD9_CODES_LIST)
```

Creating Dummy variables
(https://stackoverflow.com/questions/56264462/how-to-convert-comma-separated-multiple-responses-into-dummy-coded-columns-in-r/56264964#56264964)

```{r}
icd.split <- strsplit(df_diag_C$ICD9_CODES_LIST, ",")
lev <- unique(unlist(icd.split))
icd.dummy <- lapply(icd.split, function(x) table(factor(x, levels=lev)))

df_diag_D <- with(df_diag_C, data.frame(HADM_ID, SUBJECT_ID, do.call(rbind, icd.dummy)))
```

```{r}
setwd("~/Documents/_DATA/subset/")
write_csv(df_diag_D, "df_diag_5.csv")
```

