---
title: "Tidymodels Test 3"
output: "html_notebook"
---

Author: Roberto Williams Batista
Email: rbatista7484@floridapoly.edu

*Notes:* 

Using variables: ETHNICITY_MAPPED, Respiratory, and LOS_GROUP_ICU.
Trimmed LOS > 90 days
This notebook is the adaptation of the classification notebook created.

---

Library

---

```{r library,warning = FALSE, messages = FALSE}
library("magrittr")
library("readr")
library("dplyr")
library("ggplot2")
library("tibble")
library("tidymodels")
library("skimr")

set.seed(2020)
```


Loading the datasets

---

```{r loading data, warning=FALSE, message=TRUE}
setwd("~/OTHERS/THESIS-II/thesis/NOTEBOOKS/_MODELS/Regression")
DATA <- read.csv("mimic_c.csv")
mimic <- DATA
```

```{r skim, eval=FALSE}
mimic %>% 
  skim()
```


```{r colnames}
mimic %>% colnames()
```

```{r Admission and 1st Care Unit, eval = FALSE}
mimic %<>%
  filter(ADMISSION_TYPE == "Urgency") %>% 
  filter(FIRST_CAREUNIT == "SICU")
```


```{r Variables Selected, eval = FALSE}
mimic %<>% 
  select(
    ETHNICITY_MAPPED,
    Respiratory,
    LOS_GROUP_ICU
  )
```


Sampling the mimic dataframe

```{r split}
mimic_split <- 
  initial_split(data = mimic, prop = 0.80)

train <- training(mimic_split)
test  <- testing(mimic_split)

```

Pre-processing with `recipes`

```{r recipe}
mimic_recipe <- 
  recipe(LOS_GROUP_ICU ~., data = train) %>% 
  #step_scale(all_predictors(), -all_outcomes()) %>%
  step_dummy(all_nominal(), -all_outcomes()) %>% 
  step_corr(all_predictors()) %>% 
  prep(retain = TRUE)

```

Executing the pre-processing
```{r executing recipe}
mimic_test <- bake(mimic_recipe, new_data = test)

mimic_train <- juice(mimic_recipe)
```

## MODELS

1. RANDOM FOREST
---

1.1 ranger
1.2 randomForest

______________________________________________________________________________________

1.1 ranger

Predictions
```{r ranger model}
mimic_ranger <- 
  rand_forest(trees = 100, mode = "regression") %>% 
  set_engine("ranger") %>% 
  fit(LOS_GROUP_ICU ~., data = mimic_train)
```

Model Validation
```{r ranger metrics}
mimic_ranger %>% 
  predict(mimic_test, type = "class") %>% 
  bind_cols(mimic_test) %>% 
  metrics(truth = LOS_GROUP_ICU, estimate = .pred_class)
```
ROC Curve
```{r ranger ROC}
mimic_ranger %>%
  predict(mimic_test, type = "prob") %>% 
  bind_cols(mimic_test) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()
```

```{r ranger ROC object}
RANGER <- mimic_ranger %>%
  predict(mimic_test, type = "prob") %>% 
  mutate(model = "Randon Forest ranger") %>% 
  bind_cols(mimic_test)
```


1.2 randomForest

Predictions
```{r random forest}
mimic_rf <- 
  rand_forest(trees = 100, mode = "classification") %>% 
  set_engine("randomForest") %>% 
  fit(LOS_GROUP_ICU ~., data = mimic_train)
```

Model Validation
```{r random forest metrics}
mimic_rf %>% 
  predict(mimic_test) %>% 
  bind_cols(mimic_test) %>% 
  metrics(truth = LOS_GROUP_ICU, estimate = .pred_class)
```
ROC Curve
```{r  random forest graphs}
mimic_rf %>%
  predict(mimic_test, type = "prob") %>% 
  bind_cols(mimic_test) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()
```

```{r random forest object}
RANDON_FOREST <- mimic_rf %>%
  predict(mimic_test, type = "prob") %>% 
  mutate(model = "Random Forest") %>% 
  bind_cols(mimic_test)
```


2. LOGISTIC REGRESSION MODEL
---

2.1 glm

Predictions
```{r logistic regression}
mimic_lreg <- 
  logistic_reg(mode  = "classification", penalty = 10, mixture = 0.1) %>%
  set_engine("glm") %>% 
  fit(LOS_GROUP_ICU ~., data = mimic_train)
```

Model Validation
```{r logistic regression metrics}
mimic_lreg %>% 
  predict(mimic_test) %>% 
  bind_cols(mimic_test) %>% 
  metrics(truth = LOS_GROUP_ICU, estimate = .pred_class)
```
ROC Curve
```{r logistic regression graphs}
mimic_lreg %>%
  predict(mimic_test, type = "prob") %>% 
  bind_cols(mimic_test) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()
```
```{r logistic regression object}
LREG <- mimic_lreg %>%
  predict(mimic_test, type = "prob") %>% 
  mutate(model = "Logist Regression") %>% 
  bind_cols(mimic_test)
```

3. BOOSTED TREES
---

3.1 xgboost

Predictions
```{r boosted tree xgboost}
mimic_boost <- 
  boost_tree(mode  = "classification", trees = 200) %>% 
  set_engine("xgboost") %>% 
  fit(LOS_GROUP_ICU ~., data = mimic_train)
```

Model Validation
```{r boost tree metrics xgboost}
mimic_boost %>% 
  predict(mimic_test) %>% 
  bind_cols(mimic_test) %>% 
  metrics(truth = LOS_GROUP_ICU, estimate = .pred_class)

```
ROC Curve
```{r boosted trees ROC graph xgboost}
mimic_boost %>%
  predict(mimic_test, type = "prob") %>% 
  bind_cols(mimic_test) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()
```

```{r boost tree xgboost object xgboost}
BT_XGBOOST <- mimic_boost %>%
  predict(mimic_test, type = "prob") %>% 
  mutate(model = "Boost Tree xgboost") %>% 
  bind_cols(mimic_test)
```

4. DECISION TREES
---

4.1 rpart

Predictions
```{r rpart decision tree}
mimic_rpart <- 
  decision_tree(mode= "classification", cost_complexity = 0.0001, min_n = 5, tree_depth = 30) %>% 
  set_engine("rpart") %>% 
  fit(LOS_GROUP_ICU ~., data = mimic_train)
```
Model Validation
```{r rpart decision tree metric}
mimic_rpart %>% 
  predict(mimic_test) %>% 
  bind_cols(mimic_test) %>% 
  metrics(truth = LOS_GROUP_ICU, estimate = .pred_class)
```
ROC Curve
```{r rpart decision tree ROC graph}
mimic_rpart %>%
  predict(mimic_test, type = "prob") %>% 
  bind_cols(mimic_test) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()
```

```{r rpart decision tree  object}
RPART_DT <- mimic_rpart %>%
  predict(mimic_test, type = "prob") %>% 
  mutate(model = "Decision Tree rpart") %>% 
  bind_cols(mimic_test)
```

4.2 c5.0 (`cost_complexity` and `tree_depthdoes` not exist in C5.0)

---

Predictions
```{r C5.0 decision tree}
mimic_c5 <- 
  decision_tree(mode  = "classification", min_n = 2) %>% 
  set_engine("C5.0") %>% 
  fit(LOS_GROUP_ICU ~., data = mimic_train)
```

Model Validation
```{r C5.0 decision tree metrics}
mimic_c5 %>% 
  predict(mimic_test) %>% 
  bind_cols(mimic_test) %>% 
  metrics(truth = LOS_GROUP_ICU, estimate = .pred_class)
```
ROC Curve
```{r C5.0 decision tree ROC graphs}
mimic_c5 %>%
  predict(mimic_test, type = "prob") %>% 
  bind_cols(mimic_test) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()
```

```{r C5 decision tree object}
C5_D_TREE <- mimic_c5 %>%
  predict(mimic_test, type = "prob") %>% 
  mutate(model = "C5.0 Decision Tree") %>% 
  bind_cols(mimic_test)
```


5. BOOST TREE VERSION

---

5.1 C5.0
Predictions
```{r C5.0 boost tree}
mimic_c5_boost <- boost_tree(mode  = "classification", min_n = 2) %>% 
  set_engine("C5.0") %>% 
  fit(LOS_GROUP_ICU ~., data = mimic_train)
```
Model Validation
```{r C5.0 boost tree metrics}
mimic_c5_boost %>% 
  predict(mimic_test) %>% 
  bind_cols(mimic_test) %>% 
  metrics(truth = LOS_GROUP_ICU, estimate = .pred_class)
```
ROC Curve
```{r C5.0 boost tree graphs}
mimic_c5_boost %>%
  predict(mimic_test, type = "prob") %>% 
  bind_cols(mimic_test) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()
```

```{r C5.0 boost tree object}
C5_BOOST <- mimic_c5_boost %>%
  predict(mimic_test, type = "prob") %>% 
  mutate(model = "C5.0 Boost Tree") %>% 
  bind_cols(mimic_test)
```

6. SVM
---

6.1 kernlab

Predictions
```{r svm}
mimic_svm <- svm_rbf(mode  = "classification") %>% 
  set_engine("kernlab") %>% 
  fit(LOS_GROUP_ICU ~., data = mimic_train)
```
Model Validation
```{r svm metrics}
mimic_svm %>% 
  predict(mimic_test) %>% 
  bind_cols(mimic_test) %>% 
  metrics(truth = LOS_GROUP_ICU, estimate = .pred_class)
```
ROC Curve
```{r svm ROC graphs}
mimic_svm %>%
  predict(mimic_test, type = "prob") %>% 
  bind_cols(mimic_test) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()
```

```{r svm object}
KLAB <- mimic_svm %>%
  predict(mimic_test, type = "prob") %>% 
  mutate(model = "SVM") %>% 
  bind_cols(mimic_test)
```

COMPARATIVE AMONG TOP 3 MODELS
---

```{r top 3 models comparison}
a<- bind_rows(RANGER, RANDON_FOREST, KLAB) %>%
  group_by(model) %>% 
  roc_curve(LOS_GROUP_ICU, .pred_A:.pred_C) %>%
  autoplot()

a+ labs(title = "Test 3",
          subtitle = "Top 3 classifiers ROC results",
          caption = "A: (LOS_ICU <= 1.8), B: (1.8 < LOS < 10), C: (LOS >= 10)")
```

Saving the notebook environment
---

```{r saving the environment image}
save.image("CLASS_test_3.RData")
```

References:
---

Kappa


https://stats.stackexchange.com/questions/82162/cohens-kappa-in-plain-english

1. Landis and Koch, who characterized values

 <0 as indicating no agreement
 0.00–0.20 as slight, 
 0.21–0.40 as fair, 
 0.41–0.60 as moderate, 
 0.61–0.80 as substantial, and 
 0.81–1 as almost perfect agreement. 

2. Fleiss's equally arbitrary guidelines characterize kappas over

">0.75 as excellent, 
 0.40 to 0.75 as fair to good, and 
<0.40 as poor."
