---
title: "Tidymodels Test"
output: "html_notebook"
#editor_options: 
#  chunk_output_type: inline
---

```{r,}
library(tidymodels)
library(skimr)
library(tibble)
library(tidyverse)
library(magrittr)
library(broom)
```

```{r}
set.seed(2020)
```

```{r}
setwd("/home/rstudio/Documents/_DATA/output/")
mimic <- read.csv("mimic_icu_d.csv")
```

```{r, eval=T}
mimic %>% summary()
```

```{r, eval=T}
mimic %>% 
  skim()
```

```{r}
mimic %>% glimpse()
```

```{r, eval=FALSE}
 mimic %<>% select( -Nervous, 
                    -Blood, 
                    -Digestive, 
                    -Place_of_occurrence, 
                    -Digestive,
                    -Skin, 
                    -Undetermined, 
                    -External,
                    -Perinatal, 
                    -Legal, 
                    -Pregnancy,
                    # -Alb_Creatine_Urine,
                    # -Creatine_Clearance,
                    # -Hematocrit,
                    # -Hematocrit_Ascites,
                    # -Hematocrit_CSF, 
                    # -Hematocrit_Joint_Fluid,   
                    # -Hematocrit_Other_Fluid, 
                    # -Hematocrit_Pleural,
                    # -Protein_Creatinine_Ratio, 
                    # -Sodium, 
                    # -Sodium_Whole_Blood, 
                    # -Uric_Acid, 
                    # -Urea_Nitrogen,
                    # -Coronary_angioplasty,
                    -HADM_ID, -LOS_GROUP_ICU,
                    #-FLAG
                    )
```

```{r}
 mimic %<>% select( #-Nervous, 
 #                    -Blood, 
 #                    -Digestive, 
 #                    -Place_of_occurrence, 
 #                    -Digestive,
 #                    -Skin, 
 #                    -Undetermined, 
 #                    -External,
 #                    -Perinatal, 
 #                    -Legal, 
 #                    -Pregnancy,
 #                    -Alb_Creatine_Urine,
 #                    -Creatine_Clearance,
 #                    -Hematocrit,
 #                    -Hematocrit_Ascites,
 #                    -Hematocrit_CSF, 
 #                    -Hematocrit_Joint_Fluid,   
 #                    -Hematocrit_Other_Fluid, 
 #                    -Hematocrit_Pleural,
 #                    -Protein_Creatinine_Ratio, 
 #                    -Sodium, 
 #                    -Sodium_Whole_Blood, 
 #                    -Uric_Acid, 
 #                    -Urea_Nitrogen,
 #                    -Coronary_angioplasty,
                     -HADM_ID, 
 #                    -LOS_GROUP_ICU,
                     #-FLAG
                    )
```


```{r}
model <- lm( formula = LOS_ICU ~ ., data = mimic, na.action = na.omit)

summary(model)
```

```{r}
mimic %>% colnames()
```

Sampling the mimic dataframe

```{r}
mimic_split <- initial_split(data = mimic, prop = 0.80)

mimic_train <- training(mimic_split)
mimic_test  <- testing(mimic_split)
```

Pre-processing with `recipes`

```{r}
mimic_recipe <- 
  recipe(LOS_ICU ~., data = mimic_train) %>%
  step_corr(all_numeric(), -all_outcomes()) %>% 
  step_center(all_numeric(), -all_outcomes()) %>% 
  step_scale(all_numeric(), -all_outcomes()) %>% 
  #step_dummy(all_nominal(), -all_outcomes()) %>% 
  prep(retain = TRUE)
```

Executing the pre-processing
```{r}
train_baked <- bake(mimic_recipe, new_data = mimic_train)
test_baked  <- bake(mimic_recipe, new_data = mimic_test)
```

```{r}
model <- lm( formula = LOS_ICU ~ ., data = mimic)
summary(model)
```


MODELS

RANDOM FOREST
---
ranger
```{r}
ranger_fit <- rand_forest(trees = 100, mode = "regression") %>% 
  set_engine("ranger") %>% 
  fit(LOS_ICU ~., data = train_baked)
```
Model Validation
```{r}
ranger_fit %>% 
  predict(test_baked) %>% 
  bind_cols(test_baked) %>% 
  metrics(truth = LOS_ICU, estimate = .pred)
```

randomForest
```{r}
rf_fit <- rand_forest(trees = 100, mode = "regression") %>% 
  set_engine("randomForest") %>% 
  fit(LOS_ICU ~., data = train_baked)
```
Model Validation
```{r}
rf_fit %>% 
  predict(test_baked) %>% 
  bind_cols(test_baked) %>% 
  metrics(truth = LOS_ICU, estimate = .pred)
```


BOOSTED TREES
---

```{r}
boost_fit <- boost_tree(trees = 200, mode = "regression" ) %>% 
  set_engine("xgboost") %>% 
  fit(LOS_ICU ~., data = train_baked)
```
Model Validation
```{r}
boost_fit %>% 
  predict(test_baked) %>% 
  bind_cols(test_baked) %>% 
  metrics(truth = LOS_ICU, estimate = .pred)
```


DECISION TREES
---
rpart
```{r}
dtrees_fit <- decision_tree(mode = "regression", cost_complexity = 0.0001, min_n = 5, tree_depth = 30) %>% 
  set_engine("rpart") %>% 
  fit(LOS_ICU ~., data = train_baked)
```
Model Validation
```{r}
dtrees_fit %>% 
  predict(test_baked) %>% 
  bind_cols(test_baked) %>% 
  metrics(truth = LOS_ICU, estimate = .pred)
```


LINEAR REGRESSION
---
lm
```{r}
lm_fit <- linear_reg(mode = "regression", penalty = NULL, mixture = NULL) %>% 
  set_engine("lm") %>% 
  fit(LOS_ICU ~., data = train_baked)
```


Model Validation
```{r}
lm_fit %>% 
  predict(test_baked) %>% 
  bind_cols(test_baked) %>% 
  metrics(truth = LOS_ICU, estimate = .pred)
```



__References:__

Kappa
https://stats.stackexchange.com/questions/82162/cohens-kappa-in-plain-english
Landis and Koch, who characterized values

 <0 as indicating no agreement
 0.00–0.20 as slight, 
 0.21–0.40 as fair, 
 0.41–0.60 as moderate, 
 0.61–0.80 as substantial, and 
 0.81–1 as almost perfect agreement. 

Fleiss's equally arbitrary guidelines characterize kappas over

">0.75 as excellent, 
 0.40 to 0.75 as fair to good, and 
<0.40 as poor."
